## Dimension reduction

```{r}
set.seed(12345)
par(mar = rep(0.2, 4))
dataMatrix <- matrix(rnorm(400), nrow=40)
image(1:10, 1:40, t(dataMatrix)[, nrow(dataMatrix):1])
heatmap(dataMatrix)
```


## What if we add a pattern ?

```{r}
set.seed(678910)
for (i in 1:40) {
    # flip a coin
    coinFlip <- rbinom(1, size = 1, prob = 0.5)
    # if coin is heads add a common pattern to that row
    if (coinFlip) {
        dataMatrix[i, ] <- dataMatrix[i, ] + rep(c(0, 3), each = 5)    
    }
}

image(1:10, 1:40, t(dataMatrix)[, nrow(dataMatrix):1])
heatmap(dataMatrix)
```

## Patterns in rows and columns

```{r}
hh <- hclust(dist(dataMatrix))
dataMatrixOrdered <- dataMatrix[hh$order, ]
par(mfrow=c(1, 3))
image(t(dataMatrixOrdered)[, nrow(dataMatrixOrdered):1])
plot(rowMeans(dataMatrixOrdered), 40:1, xlab="Row mean", ylab="Row", pch=19)
plot(colMeans(dataMatrixOrdered), xlab="Column", ylab="Column Mean", pch=19)
```

## Related problems

You have multivariate variables X1, X2, X3...Xn so X1 = (X11, X12..X1n)

*   Find a new set of multivariate variables that are uncorrelated and explain as much variance as possible
    - statistical
    - principal component analysis

* If you put all the variables together in one matrix, find the best matrix created with fewer varibles that explains the original data
    - data compression
    - singular value decomposition
    
## Related solutions - PCA/SVD

### SVD

If X is a matrix with each variable in a column and each observation in a row, then a the SVD is a "matrix decomposition"

             **X = UDV(sup T)**
             
where the columns of U are orthogonal (left singluar vectors), the columns of V are orthogonal (right singular vectors) and D is a diagonal matrix (singular values)

## PCA

The principal components are equal to the right singular values, if you first scale(subtract the mean divide by the standard deviation) the variables

## Components of the SVD - u and v

```{r}
svd1 <- svd(scale(dataMatrixOrdered))
par(mrfrow=c(1, 3), mar=rep(4, 4))
image(t(dataMatrixOrdered)[, nrow(dataMatrixOrdered):1])
plot(svd1$u[, 1], 40:1, xlab="Row", ylab="First left singular vector", pch=19)
plot(svd1$v[, 1], xlab="Column", ylab="Frist right singular vector", pch=19)
```

## Components of the SVD - Variance explained

```{r}
par(mfrow=c(1, 2))
plot(svd1$d, xlab="Column", ylab="Singular value", pch=19)
plot(svd1$d^2/sum(svd1$d^2), xlab="Column", ylab="Prop. of variance explained", pch=19)
```

## Relationship to principal components

```{r}
svd1 <- svd(scale(dataMatrixOrdered))
pca1 <- prcomp(dataMatrixOrdered, scale=TRUE)
plot(pca1$rotation[, 1], svd1$v[, 1], pch=19, xlab="Principal Component 1", ylab="Right Singular Vector 1")
abline(c(0, 1))
```

## Components of the SVD - variance explained

```{r}
constantMatrix <- dataMatrixOrdered*0
for (i in 1:dim(dataMatrixOrdered)[1]) {
    constantMatrix[i, ] = rep(c(0, 1), each=5)
}
svd1 <- svd(constantMatrix)
par(mfrow = c(1, 3))
image(t(constantMatrix)[, nrow(constantMatrix):1])
plot(svd1$d, xlab="Column", ylab="Singular value", pch=19)
plot(svd1$d^2/sum(svd1$d^2), xlab="Column", ylab="Prop. of variance explained", pch=19)
```

## What if we add a second pattern ?

```{r}
set.seed(678910)
for (i in 1:40) {
    #flip a coin
    coinFlip1 <- rbinom(1, size=1, prob=0.5)
    coinFlip2 <- rbinom(1, size=1, prob=0.5)
    if (coinFlip1) {
        dataMatrix[i, ] <- dataMatrix[i, ] + rep(c(0, 5), each=5)
    }
    if (coinFlip2) {
        dataMatrix[i, ] <- dataMatrix[i, ] + rep(c(0, 5), 5)
    }
}
hh <- hclust(dist(dataMatrix))
dataMatrixOrdered <- dataMatrix[hh$order, ]
```

## Singular value decomposition - true patterns

```{r}
svd2 <- svd(scale(dataMatrixOrdered))
par(mfrow = c(1, 3))
image(t(dataMatrixOrdered)[, nrow(dataMatrixOrdered):1])
plot(rep(c(0, 1), each=5), pch=19, xlab="Column", ylab="Pattern 1")
plot(rep(c(0, 1), 5), pch=19, xlab="Column", ylab="Pattern 2")
```

## v and patterns of variance in rows

```{r}
svd2 <- svd(scale(dataMatrixOrdered))
par(mfrow=c(1, 3))
image(t(dataMatrixOrdered)[, nrow(dataMatrixOrdered):1])
plot(svd2$v[, 1], pch=19, xlab="Column", ylab="First right singular vector")
plot(svd2$v[, 2], pch=19, xlab="Column", ylab="Second right singular vector")
```

## d and variance explained

```{r}
svd1 <- svd(scale(dataMatrixOrdered))
par(mfrow=c(1, 2))
plot(svd1$d, xlab="Column", ylab="Singular value", pch=19)
plot(svd1$d^2/sum(svd1$d^2), xlab="Column", ylab="Percent of variance explained", pch=19)
```