# Hierarachical Clustering

* Clustering organizes things that are close into groups

Different clustering algorithms are variations on

* How do we define close ?
* How do we group things ?
* How do we visualize the grouping ?
* How do we interpret the group ?

* An agglomerative Approaches
    * Find closest two things
    * put them together (create a new data point, merge of the above two points)
    * find next closest

* Requires
    
    * a defined distance
    * a merging approach
    
* Produces
    * a tree showing how close things are to each other
    
### How do we define close ?

* Most important step
    - Garbage in -> Garbage out

* Distance or similarity
    - Continuous - euclidean distance (straight line distance)
    - Continuous - correlation similarity
    - Binary - manhattan distance
    
* Pick a distance/similarity that makes sense for that problem

### Euclidean

* euclidean distance -> sqrt(sqr(x2-x1) + sqr(y2-y1))
* bird flight distance
* easy for higher dimensional problems

### Manhattan 
* abs(x2-x1) + abs(y2 - y1) + ... + abs(z2 - z1)

## Hierarchical clustering - example

````{r}
set.seed(1234)
par(mar=c(0, 0, 0, 0))
x <- rnorm(12, mean=rep(1:3, each=4), sd = 0.2)
y <- rnorm(12, mean=rep(c(1,2,1), each=4), sd = 0.2)
plot(x, y, col="blue", pch=19, cex=2)
text(x + 0.05, y + 0.05, labels=as.character(1:12))
```

```{r}
dataFrame <- data.frame(x = x, y = y)
distxy <- dist(dataFrame) # takes first two columns, defaults to euclidean distance
```

### Hierarchical clustering using hclust
```{r}
hClustering <- hclust(distxy)
par(mar=c(4, 4, 4, 4))
plot(hClustering)
```

```{r}
source("./myplclust.R")
myplclust(hClustering, lab=rep(1:3, each=4), lab.col=rep(1:3, each=4))
```

## Mergin points
 
* average linkage - average of x,y coordinates of two clusters
* complete linkage - average of two furthest points in the clusters

```{r}
set.seed(143)
dataMatrix <- as.matrix(dataFrame)[sample(1:12),]
heatmap(dataMatrix)
```

## Notes

* gives an idea of the relationships between variables/observations

* The picture may be unstable
    - Change a few points
    - Have different missing values
    - Pick a different strategy
    - Change the merging strategy
    - Change the scale of points for one variable
    
* But it is deterministic
* Choosing where to cut isn't always obvious
* Should be primarily used for exploration