## K-means clustering

* an approach
    - fixed number of clusters (starting with k random centers)
    - assign each point to the closest centroid (some distance measruing algorithm)
    - get "centroids" of each new cluster
    - recalculate "centroids"  (step 2)
    - till there is no jumping of "points" to different cluster
    
* requires
    - A defined distance metric
    - A number of clusters
    - An initial guess as to cluster centroids
    
* produces
    - final estimate of cluster centroids
    - An assignment of each point to clusters
    
```{r}
set.seed(1234)
par(mar=c(0, 0, 0, 0))
x <- rnorm(12, mean=rep(1:3, each=4), sd = 0.2)
y <- rnorm(12, mean=rep(c(1,2,1), each=4), sd = 0.2)
plot(x, y, col="blue", pch=19, cex=2)
text(x + 0.05, y + 0.05, labels=as.character(1:12))
```

```{r}
dataFrame <- data.frame(x =x, y = y)
kmeansObj <- kmeans(dataFrame, centers=3)
names(kmeansObj)
kmeansObj$cluster
```

```{r}
par(mar=rep(0.2, 4))
plot(x, y, col = kmeansObj$cluster, pch=19, cex=2)
points(kmeansObj$centers, col = 1:3, pch=3, cex = 3, lwd=3)
```

## Heatmaps

```{r}
set.seed(1234)
dataMatrix <- as.matrix(dataFrame)[sample(1:12), ]
kmeansObj2 <- kmeans(dataMatrix, centers=3)
par(mfrow = c(1, 2), mar=c(2, 4, 0.1, 0.1))
image(t(dataMatrix)[, nrow(dataMatrix):1], yaxt = "n")
image(t(dataMatrix)[, order(kmeansObj$cluster)], yaxt="n")
```

## Notes and further resources

* K - means requires a number of clusters
    - Pick by eye/intuition
    - determin the number of clusters
* Kmean is not deterministic
    - Different # of clusters
    - Diffeernt number of iterations