## Plotting and Color

* the default color schemes for most plots in R are bad !
* There are functions in R and in external packages that are very handy

* colors() - list of color names
* colorRamp() - takes a palette of colors and returns a function(0 to 1) 
    - gray is an example - with palette being white and black
* colorRampPalette() - take a palette of colors and return a function that takes integer arguments and returns a vector of colors interpolating the palette 
    - heat.colors, topo.colors
    
```{r}
pal <- colorRamp(c("red", "blue"))
pal(0)
pal(0.5)
pal(1)
pal(seq(0, 1, len=10))
```

```{r}
pal <- colorRampPalette(c("red", "yellow", "black"))
pal(2)
pal(30)
```

## RColorBrewer Package

* One package on CRAN that contains useful color palettes
* 3 palettes
    - Sequential
    - Diverging (showing deviation from mean etc..)
    - Qualitative
* Palette information can be used in conjunction with colorRamp(), colorRampPalette()

```{r}
library(RColorBrewer)
cols <- brewer.pal(3, "BuGn")
cols
pal <- colorRampPalette(cols)
image(volcano, col=pal(20))
```

```{r}
x <- rnorm(10000)
y <- rnorm(10000)
smoothScatter(x, y) # reflects density of distribution
```

## Some other plotting notes

* rgb - function decimal to hexdecimal
* alpha param for rgb
* the colorspace package can be used for a different control over colors


## Scatter plot with transparency
```{r}
par(mfrow=c(1,2))
plot(x, y, pch=19) # blotchy
plot(x, y, col=rgb(0, 0, 0, 0.2), pch=19)
```