My First knitr document
=====================

```{r setoptions, echo=TRUE, results="asis", fig.height=4, fig.width=4}
opts_chunk$set(echo = FALSE, results="hide")
```

This is some text (i.e a "text chuck")

Here is a code chunk

```{r}
set.seed(1)
x <- rnorm(100)
mean(x)
```

# My First knitr doc
Chandra Koduru

## Introduction

```{r simulation, echo=FALSE, results="hide"}
set.seed(1)
x <- rnorm(100)
mean(x)
```

```{r computetime, echo=FALSE}
time <- format(Sys.time(), "%a %b %d %X %Y")
rand <- rnorm(1)
```

The current time is **`r time`**. My favorite random number is `r rand`.

```{r simulatedata, echo=TRUE}
x <- rnorm(100)
y <- x + rnorm(100, sd = 0.5)
```

Here is a scatter plot of the data
```{r scatterplot, fig.height=4}
par(mar = c(5,4,1,1), las=1)
plot(x, y, main="My Simulated Data", xlab="x label")
```

* embedded image is base 64 encoded

## Presenting tabular data

```{r fitmodel, echo=TRUE}
library(datasets)
data(airquality)
fit <- lm(Ozone ~ Wind + Temp + Solar.R, data = airquality)
```

Here is a table of regression coefficients.
```{r showtable, results="asis"}
library(xtable)
xt <- xtable(summary(fit))
print(xt, type="html")
```

## Setting global options
* for every code chunk
* such as supressing echo

## Cachcing Computations

* set 'cache=TRUE' option can be set on chunk-by-chunk basis to store results of computation